---
title: "R4MS Workshop 2 - Final Exercise"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
```

# Introduction

Here we are going to wrangling in messy data on the Ross River estuary fish surveys and produce final clean tables and figure. We have 4 separate files to wrangle in and use

# Housekeeping

```{r housekeeping}
#| message: true
#| warning: false
rm(list = ls())
```

# Packages

```{r}
library(tidyverse) # Inc dplyr, ggplot2, lubridate etc
library(readxl) # For reading in excel files
```

# Load files

# Main dataset

The catch data is split across 4 tabs in one excel file Load indivudally

```{r}
catch1 <- read_excel(here::here("data/workshop2/estuary_catch_log.xlsx"), sheet = 1)
catch2 <- read_excel("../data/workshop2/estuary_catch_log.xlsx", sheet = 2)
catch3 <- read_excel("../data/workshop2/estuary_catch_log.xlsx", sheet = 3)
catch4 <- read_excel("../data/workshop2/estuary_catch_log.xlsx", sheet = 4)

catch_data <- bind_rows(catch1, catch2, catch3, catch4)
# Clean up the environment
rm(catch1, catch2, catch3, catch4) # Clean up the environment
```

# Alternative method for loading files

```{r}
# file_path <- "../data/workshop2/estuary_catch_log.xlsx"
# catch_data2 <- excel_sheets(file_path) |>
#   map_dfr(~read_excel(path = file_path, sheet = .x))
```

# Examine data

```{r}
catch_data
head(catch_data)
glimpse(catch_data)
str(catch_data)
```

Can't tell from these outputs how many levels of site and species we have. Two options: let's check those explicitly by using distinct() or can convert to factors Let's look at the distinct values first, as we might need to clean before converting to factor vars.

```{r}
catch_raw |> 
  distinct(site)

catch_raw |> 
  distinct(species)
```
